# Tooth Growth Data Analysis
authored by Gran Ville Lintao
June 22, 2015

## Overview
This document contains statistical analysis for the ToothGrowth dataset in R.  
It is divided into 4 parts:  
1. __Exploratory data analysis__ summarizing an initial observation  
2. Usage of __Confidence Intervals__ for the mean difference of the two types of supplement  
3. Usage of a Hypothesis Test - __Unequal Variance T Test__  for the two types of supplement  
4. __Assumptions and Conclusions__ about the tests performed.

## Exploratory Data Analysis

In this part we load the libraries needed, the data, and then using plyr, _generate a summary_ by 
averaging the toothGrowth by supplement and dose:
```{r, echo=TRUE}
library(ggplot2)
library(plyr)

data(ToothGrowth)
dataSum <- ddply(ToothGrowth, ~supp * dose, summarise, 
                 aveGrowth=ave(len))
```

We then plot this summary and we can quickly see the positive correlation between dose and length of tooth growth
on both doses. Furthermore, we can see that the supp __OJ__ is correlated with __higher tooth growth__ compared to __VC__
on the doses __0.5 and 1.0__.

```{r, echo=TRUE}
qplot(supp, aveGrowth, data=dataSum, facets=.~dose,
      main="Average Tooth Growth by Dose (milligrams)",
      ylab="Average Tooth Growth",
      xlab="Supplement Type")
```


## Usage of Confidence Intervals
In this part, we use confidence intervals to compare the means between the doses VC and OJ.
Note that these are independent groups. 

First we take divide them in g1 and g2 for easier computing
```{r, echo=TRUE}
g1 <- subset(ToothGrowth, supp=="VC")$len
g2 <- subset(ToothGrowth, supp=="OJ")$len
n1 <- length(g1); n2 <- length(g2)
```

Then we use can use the following computation for manually getting the confidence intervals
```{r, echo=TRUE}
# confidence intervals option 1 manual computation practice :)
sp <- sqrt( ((n1-1) * sd(g1)^2 + (n2-1) * sd(g2)^2) / (n1+n2-2) ) # pooled sd estimate
md <- mean(g2) - mean(g1) # mean difference
semd <- sp * sqrt(1/n1 + 1/n2) # std error of the mean difference
ciManual <- md + c(-1,1) * qt(.975,n1+n2-2) * semd
ciManual
```

We can also use the t.test() function to check if our computation above is correct
```{r, echo=TRUE}
# confidence intervals option 2 using t.test
ciTTest <- t.test(g2, g1, paired=FALSE, var.equal=TRUE)$conf
ciTTest
```

Here we can see that the manually computed __ciManual__ : `r ciManual` is the same with __ciTTest__ : `r ciTTest`


## Unequal Variance T Test
Here we use an unequal variance T Test for comparing whether the means in the 
two supplements OJ and VC are equal under the null Hypothesis. The alternative
hypothesis is that they're different.

```{r, echo=TRUE}
uvTTest <- t.test(data=ToothGrowth, len ~ supp, paired=FALSE, var.equal=TRUE)$statistic
uvTTest
```

Here we can see that the estimated standard errors the difference in means from the hypothesized mean is : `r uvTTest`.
Based on this we reject the null Hypothesis.

## Assumptions and Conclusions

__Conclusions:__  
1. Larger dosages is correlated with larger tooth growth  
2. dose OJ has larger toothgrowth in lower doses 0.5, 1.0 compared to dose VC  
3. We use an Indenpendent Group T Confidence Intervals due to assumption 1  
4. The means in supplements OJ and VC are not equal - We reject the null hypothesis in favor of the alternative  

__Assumptions:__  
1. Groups OJ and VC are independent and not paired  
2. We use an Unequal Variance T Test for comparing OJ and VC, which can be treated as 'diets'  



